---
title: "Equivalence tests with scaled variances"
geometry: margin=0.5in
output:
  pdf_document:
    df_print: kable
    toc: yes
    toc_depth: 4
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage

# Preparations

## Load packages

```{r message=FALSE, warning=FALSE}
library(psych)
library(lavaan)
library(dplyr)
```


## Import data

```{r}
dat<-read.csv2("dat.no.miss.csv",stringsAsFactors = F)
```

\newpage

# Equivalence test for Attitudes towards Refugees'

## Data

```{r}
#exclude Hungary
dat.imm<-dat %>%
  filter(cntry!="HU") %>%
  select(cntry,gvrfgap.R,rfgfrpc,rfgbfml.R)
dat.imm<-na.omit(dat.imm)

```

## Model script

```{r}
imm.mod<-'
#loadings
imm.fac=~gvrfgap.R+rfgfrpc+rfgbfml.R

#latent variance
imm.fac~~c(v1,v2,v3,v4,v5,v6,v7,v8,v9,v10,v11,v12,v13,v14,v15,v16,v17,v18,v19)*imm.fac
#latent mean
#imm.fac~c(m1,m2,m3,m4,m5,m6,m7,m8,m9,m10,m11,m12,m13,m14,m15,m16,m17,m18,m19)*1

#observed means
gvrfgap.R~1
rfgfrpc~1
rfgbfml.R~1

#observed variances
gvrfgap.R~~gvrfgap.R
rfgfrpc~~rfgfrpc
rfgbfml.R~~rfgbfml.R

#constrain sds
sqrt(v1*v2*v3*v4*v5*v6*v7*v8*v9*v10*v11*v12*v13*v14*v15*v16*v17*v18*v19)==1

'
```


## Fit Multi-group confirmatory factor analysis

```{r}
metric.imm<-lavaan(model=imm.mod,
                   data=dat.imm,
                   group="cntry",
              group.equal=c("loadings"),
              estimator="ML")

```

\newpage

## Print the model output

```{r}
summary(metric.imm,fit=T,rsquare=T)
```

\newpage

## Equivalence tests

```{r paged.print=TRUE}

#obtain the tests for all constrained parameters
imm.eq.tests<-lavTestScore(metric.imm,
                           epc=T,standardized=T)

#exclude non-specified parameters
imm.eq.tests$epc<-imm.eq.tests$epc[imm.eq.tests$epc$label!="" &
                                     grepl("p",imm.eq.tests$epc$label),]

#combine the univariate test and epc frames
imm.eq.tests<-left_join(x=imm.eq.tests$epc,
                        y=imm.eq.tests$uni,
                        by=c("plabel"="rhs"))

#get country names
groups.imm<-cbind.data.frame(group=1:19,cntry=inspect(metric.imm,"group.label"))
#combine to equivalence test data
imm.eq.tests<-left_join(x=imm.eq.tests,
                        y=groups.imm,
                        by=c("group"))

#define the preregistered criteria for decisions

#maximum absolute EPC
delta=.10
#power limit
high.power=.80
#type-I error
alpha=.05

#calculate power based with non-centrality parameters

imm.eq.tests$ncp<-
  (delta^2)*imm.eq.tests$X2/(imm.eq.tests$epc^2)
imm.eq.tests$power<-
  pchisq(ncp = imm.eq.tests$ncp,
         df=1,q=qchisq(p=alpha,df=1,lower.tail=F),lower.tail=F)



#code the decisions (according to Saris et al., 2009)

imm.eq.tests$decision<-
  ifelse(
    (imm.eq.tests$power>=high.power &
       imm.eq.tests$p.value >=alpha),"nm",
    ifelse(
      (imm.eq.tests$power>=high.power &
         imm.eq.tests$p.value <alpha &
         abs(imm.eq.tests$epc) <=delta),"epc:nm",
      ifelse(
        (imm.eq.tests$power>=high.power &
           imm.eq.tests$p.value <alpha &
           abs(imm.eq.tests$epc) >delta),"misspec",
        ifelse(
          (imm.eq.tests$power<high.power & imm.eq.tests$p.value <alpha),"misspec",
          ifelse(
            (imm.eq.tests$power<high.power & imm.eq.tests$p.value >alpha),"inconcl",
            "")))))

#summary of the decisions across all the parameters
table(imm.eq.tests$decision)

#Table that only contains the important numbers
imm.eq.tests.tab<-cbind(imm.eq.tests[,c("rhs","cntry","decision")],
                        round(
                          imm.eq.tests[,c("est","epc","epv",
                                          "sepc.all","X2","power","p.value")],2))
#print the table
imm.eq.tests.tab
#save to drive
write.csv2(imm.eq.tests.tab,"imm.eq.scaled.tests.tab.csv")

#country-specific decisions
table(imm.eq.tests.tab$cntry,
      imm.eq.tests.tab$decision)


#look for only epc-based tests for .10, .15, and .20

table(imm.eq.tests$cntry,
      abs(imm.eq.tests$epc)>.10)


table(imm.eq.tests$cntry,
      abs(imm.eq.tests$epc)>.15)


table(imm.eq.tests$cntry,
      abs(imm.eq.tests$epc)>.20)

```

\newpage

## Summary for Equivalence test for Attitudes towards Refugees'

* Of the 54 parameter constrains (19 countries * 3 loadings - 1 country as a reference group), 22 were considered as misspecification, 18 were not misspecifications (nm or epc:nm), and 14 were inclonclusive (inconcl) due to low power for statistical tests
  + These misspecifications were spread across countries so that 7 countries had more than one misspecification
  + This would lead to using the single item scale for measuring Attitudes towards Refugees' by the preregistered criteria
* Observations based on only EPC (expected parameter change) showed that the loadings of the reference group (Austria) were not misspecified (epc >|.10|)
  + In total 10 countries would have had 2 or more misspecifications based only on EPCs (.10)
  + If the criterion would have been .15 instead of .10, 3 countries would have been non-equivalent (2 or 3 misspecifications)
  + If the criterion would have been .20 instead of .10, only Czech and Lithuania would have been non-equivalent
  


\newpage

# Equivalence test for Attitudes towards the Environment

## Data

```{r}

dat.env<-dat %>%
  select(cntry,inctxff.R,sbsrnen.R,banhhap.R)
dat.env<-na.omit(dat.env)

```

## Model script

```{r}
env.mod<-'

#loadings
env.fac=~inctxff.R+sbsrnen.R+banhhap.R

#latent variance
env.fac~~c(v1,v2,v3,v4,v5,v6,v7,v8,v9,v10,v11,v12,v13,v14,v15,v16,v17,v18,v19,v20)*env.fac
#latent mean
#env.fac~1

#observed means
inctxff.R~1
sbsrnen.R~1
banhhap.R~1

#observed variances
inctxff.R~~inctxff.R
sbsrnen.R~~sbsrnen.R
banhhap.R~~banhhap.R

#constrain sds
sqrt(v1*v2*v3*v4*v5*v6*v7*v8*v9*v10*v11*v12*v13*v14*v15*v16*v17*v18*v19*v20)==1

'
```

## Fit Multi-group confirmatory factor analysis

```{r}

metric.env<-lavaan(model=env.mod,data=dat.env,group="cntry",
              group.equal=c("loadings"),estimator="ML")

```

\newpage

## Print the model output

```{r}
summary(metric.env,fit=T,rsquare=T)
```

\newpage

## Equivalence tests

```{r paged.print=TRUE}


#obtain the tests for all constrained parameters
env.eq.tests<-lavTestScore(metric.env,
                           epc=T,standardized=T)

#exclude non-specified parameters
env.eq.tests$epc<-env.eq.tests$epc[env.eq.tests$epc$label!="" &
                                  grepl("p",env.eq.tests$epc$label),]


#combine the univariate test and epc frames
env.eq.tests<-left_join(x=env.eq.tests$epc,
                        y=env.eq.tests$uni,
                        by=c("plabel"="rhs"))

#get country names
groups.env<-cbind.data.frame(group=1:20,cntry=inspect(metric.env,"group.label"))
#combine to equivalence test data
env.eq.tests<-left_join(x=env.eq.tests,
                        y=groups.env,
                        by=c("group"))

#define the preregistered criteria for decisions

#maximum absolute EPC
delta=.10
#power limit
high.power=.80
#type-I error
alpha=.05

#calculate power based with non-centrality parameters

env.eq.tests$ncp<-
  (delta^2)*env.eq.tests$X2/(env.eq.tests$epc^2)
env.eq.tests$power<-
  pchisq(ncp = env.eq.tests$ncp,
         df=1,q=qchisq(p=alpha,df=1,lower.tail=F),lower.tail=F)



#code the decisions (according to Saris et al., 2009)

env.eq.tests$decision<-
  ifelse(
    (env.eq.tests$power>=high.power &
       env.eq.tests$p.value >=alpha),"nm",
    ifelse(
      (env.eq.tests$power>=high.power &
         env.eq.tests$p.value <alpha &
         abs(env.eq.tests$epc) <=delta),"epc:nm",
      ifelse(
        (env.eq.tests$power>=high.power &
           env.eq.tests$p.value <alpha &
           abs(env.eq.tests$epc) >delta),"misspec",
        ifelse(
          (env.eq.tests$power<high.power & env.eq.tests$p.value <alpha),"misspec",
          ifelse(
            (env.eq.tests$power<high.power & env.eq.tests$p.value >alpha),"inconcl",
            "")))))

#summary of the decisions across all the parameters
table(env.eq.tests$decision)

#Table that only contains the important numbers
env.eq.tests.tab<-cbind(env.eq.tests[,c("rhs","cntry","decision")],
                        round(
                          env.eq.tests[,c("est","epc","epv",
                                          "sepc.all","X2","power","p.value")],2))
#print the table
env.eq.tests.tab
#save to drive
write.csv2(env.eq.tests.tab,"env.eq.scaled.tests.tab.csv")

#country-specific decisions
table(env.eq.tests.tab$cntry,
      env.eq.tests.tab$decision)


#look for only epc-based tests for .10, .15, and .20

table(env.eq.tests$cntry,
      abs(env.eq.tests$epc)>.10)


table(env.eq.tests$cntry,
      abs(env.eq.tests$epc)>.15)


table(env.eq.tests$cntry,
      abs(env.eq.tests$epc)>.20)
```


\newpage

## Summary for Equivalence test for Attitudes towards the Environment

* Of the 57 parameter constrains (20 countries * 3 loadings - 1 country as a reference group), 26 were considered as misspecifications, 22 were not misspecifications (nm or epc:nm), and 9 were inclonclusive (inconcl) due to low power for statistical tests
  + These misspecifications were spread across countries so that 9 countries had more than one misspecification
  + This would lead to using the single item scale for measuring Attitudes towards the Environment by the preregistered criteria
* Observations based on only EPC (expected parameter change) showed that the loadings of the reference group (Austria) were not misspecified (epc <|.10|)
  + In total 9 countries would have had 2 or more misspecifications based only on EPCs (.10)
  + If the criterion would have been .15 instead of .10, 5 countries would have been non-equivalent (2 or 3 misspecifications)
  + If the criterion would have been .20 instead of .10, only Hungary and Poland would have been non-equivalent

\newpage

# Session and package information

```{r}
sessionInfo()
```
