---
title: "Exlopratory Hypothesis 5 with Political News Consumption"
output: 
  html_document: 
    keep_md: yes
    toc: yes
    toc_depth: 5
---

# Exploratory Hypothesis 5: The strength of the association between environment and refugee attitudes is stronger among those who consume political news more

Omit missing values on political news consumption item

```{r}
dat.H5.news<-dat %>%
  filter(!is.na(polnews))
```

### Model 1: without interactions (only main effects)

```{r}
H5.exp.news.mod1<-lmer(refugees~(environ.lvl1|voting.group)+
                (environ.lvl1|cntry)+
                age+gender+educ+resid+occup+
                environ.lvl1+
                polnews.lvl1
                ,data=dat.H5.news,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))


isSingular(H5.exp.news.mod1)
(FE.H5.exp.news.mod1<-getFE(H5.exp.news.mod1))
(VC.H5.exp.news.mod1<-getVC(H5.exp.news.mod1))


```

\newpage

### Model 2: Level-1 interaction between environmental attitudes and political polnews

```{r}
H5.exp.news.mod2<-lmer(refugees~(environ.lvl1|voting.group)+
                (environ.lvl1|cntry)+
                age+gender+educ+resid+occup+
                environ.lvl1+
                polnews.lvl1+
                environ.lvl1:polnews.lvl1
                ,data=dat.H5.news,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))
isSingular(H5.exp.news.mod2)
anova(H5.exp.news.mod1,H5.exp.news.mod2)

(FE.H5.exp.news.mod2<-getFE(H5.exp.news.mod2))
(VC.H5.exp.news.mod2<-getVC(H5.exp.news.mod2))

```

\newpage

### Model 3: Level-1 interaction between environmental attitudes and political polnews, allow polnews effect to vary between voting groups and countries

```{r}
H5.exp.news.mod3<-lmer(refugees~(environ.lvl1+polnews.lvl1|voting.group)+
                (environ.lvl1+polnews.lvl1|cntry)+
                age+gender+educ+resid+occup+
                environ.lvl1+
                polnews.lvl1+
                environ.lvl1:polnews.lvl1
                ,data=dat.H5.news,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))
isSingular(H5.exp.news.mod3)
anova(H5.exp.news.mod2,H5.exp.news.mod3)

(FE.H5.exp.news.mod3<-getFE(H5.exp.news.mod3))
(VC.H5.exp.news.mod3<-getVC(H5.exp.news.mod3))

```


\newpage

### Model 4: Allow the level-1 interaction to vary between voting groups and countries

```{r}
H5.exp.news.mod4<-lmer(refugees~
                (environ.lvl1+polnews.lvl1+
                   environ.lvl1:polnews.lvl1|voting.group)+
                (environ.lvl1+polnews.lvl1+
                   environ.lvl1:polnews.lvl1|cntry)+
                age+gender+educ+resid+occup+
                environ.lvl1+
                polnews.lvl1+
                environ.lvl1:polnews.lvl1
                ,data=dat.H5.news,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))
isSingular(H5.exp.news.mod4)
anova(H5.exp.news.mod3,H5.exp.news.mod4)

(FE.H5.exp.news.mod4<-getFE(H5.exp.news.mod4))
(VC.H5.exp.news.mod4<-getVC(H5.exp.news.mod4))


theta <- getME(H5.exp.news.mod4,"theta")

## diagonal elements are identifiable because they are fitted
##  with a lower bound of zero ...
diag.element <- getME(H5.exp.news.mod4,"lower")==0
any(theta[diag.element]<1e-5)
round(theta,5)

```

\newpage

### Model 5: Enter voting group variable and both two-way interactions with environment and polnews

```{r}
H5.exp.news.mod5<-lmer(refugees~
                (environ.lvl1+polnews.lvl1+environ.lvl1:polnews.lvl1|voting.group)+
                (environ.lvl1+polnews.lvl1+environ.lvl1:polnews.lvl1|cntry)+
                age+gender+educ+resid+occup+
                environ.lvl1+
                polnews.lvl1+
                environ.lvl1:polnews.lvl1+
                all.parties.lvl2+
                environ.lvl1:all.parties.lvl2+
                polnews.lvl1:all.parties.lvl2
                ,data=dat.H5.news,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))
isSingular(H5.exp.news.mod5)
anova(H5.exp.news.mod4,H5.exp.news.mod5)

(FE.H5.exp.news.mod5<-getFE(H5.exp.news.mod5))
(VC.H5.exp.news.mod5<-getVC(H5.exp.news.mod5))

```

\newpage

#### Look among which voting group there is strongest association between polnews and refugee attitudes

```{r}
H5.exp.news.mod5.trends<-emtrends(H5.exp.news.mod5,specs = c("all.parties.lvl2"),var=c("polnews.lvl1"))
(H5.exp.news.mod5.trends.tab<-data.frame(H5.exp.news.mod5.trends))


H5.exp.news.mod5.trends.tab$p<-
  2*(1-pnorm(abs(H5.exp.news.mod5.trends.tab$polnews.lvl1.trend/
                   H5.exp.news.mod5.trends.tab$SE)))
H5.exp.news.mod5.trends.tab$adj.p<-
  p.adjust(H5.exp.news.mod5.trends.tab$p,method="holm")

H5.exp.news.mod5.trends.tab<-
  cbind(group=H5.exp.news.mod5.trends.tab[,1],
      round(H5.exp.news.mod5.trends.tab[,c(2,3)],2),
      round(H5.exp.news.mod5.trends.tab[,c(7,8)],4),
      round(H5.exp.news.mod5.trends.tab[,c(5,6)],2))
H5.exp.news.mod5.trends.tab

write.csv2(H5.exp.news.mod5.trends.tab,"H5.exp.news.mod5.trends.tab.csv")

#contrast between anti-immigration and pro-environment
(H5.exp.news.contrast<-data.frame(pairs(H5.exp.news.mod5.trends, exclude = c(2:9),reverse=T)))
#contrast for all groups against mean of other groups
contrast(H5.exp.news.mod5.trends, "del.eff", by = NULL,adjust=c("none"))
#contrast for three voting groups
(H4.more.contrasts<-data.frame(pairs(H5.exp.news.mod5.trends, 
         exclude=c(2:8), by = NULL,adjust=c("none"),reverse=T)))

```

\newpage

### Model 6: Enter three-way interaction voting group x polnews x environment attitudes

```{r}
H5.exp.news.mod6<-lmer(refugees~
                (environ.lvl1+polnews.lvl1+environ.lvl1:polnews.lvl1|voting.group)+
                (environ.lvl1+polnews.lvl1+environ.lvl1:polnews.lvl1|cntry)+
                age+gender+educ+resid+occup+
                environ.lvl1+
                polnews.lvl1+
                environ.lvl1:polnews.lvl1+
                all.parties.lvl2+
                environ.lvl1:all.parties.lvl2+
                polnews.lvl1:all.parties.lvl2+
                environ.lvl1:polnews.lvl1:all.parties.lvl2
                ,data=dat.H5.news,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))
isSingular(H5.exp.news.mod6)
anova(H5.exp.news.mod5,H5.exp.news.mod6)

(FE.H5.exp.news.mod6<-getFE(H5.exp.news.mod6))
(VC.H5.exp.news.mod6<-getVC(H5.exp.news.mod6))

```

#### Refit with manually coded level-1 interaction

```{r}
dat.H5.news$env.news.int<-dat.H5.news$environ.lvl1*dat.H5.news$polnews.lvl1

H5.exp.news.mod6<-lmer(refugees~
                (environ.lvl1+polnews.lvl1+env.news.int|voting.group)+
                (environ.lvl1+polnews.lvl1+env.news.int|cntry)+
                age+gender+educ+resid+occup+
                environ.lvl1+
                polnews.lvl1+
                env.news.int+
                all.parties.lvl2+
                environ.lvl1:all.parties.lvl2+
                polnews.lvl1:all.parties.lvl2+
                env.news.int:all.parties.lvl2
                ,data=dat.H5.news,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))
isSingular(H5.exp.news.mod6)
anova(H5.exp.news.mod5,H5.exp.news.mod6)

(FE.H5.exp.news.mod6<-getFE(H5.exp.news.mod6))
(VC.H5.exp.news.mod6<-getVC(H5.exp.news.mod6))

```


#### Marginal effect for pro-environment and anti-immigration voters

```{r message=FALSE}
H5.exp.news.mod6.trends<-emtrends(H5.exp.news.mod6,specs = c("all.parties.lvl2"),var=c("env.news.int"))
(H5.exp.news.mod6.trends.tab<-data.frame(H5.exp.news.mod6.trends))


H5.exp.news.mod6.trends.tab$p<-
  2*(1-pnorm(abs(H5.exp.news.mod6.trends.tab$env.news.int.trend/
                   H5.exp.news.mod6.trends.tab$SE)))
H5.exp.news.mod6.trends.tab$adj.p<-
  p.adjust(H5.exp.news.mod6.trends.tab$p,method="holm")

H5.exp.news.mod6.trends.tab<-
  cbind(group=H5.exp.news.mod6.trends.tab[,1],
      round(H5.exp.news.mod6.trends.tab[,c(2,3)],2),
      round(H5.exp.news.mod6.trends.tab[,c(7,8)],4),
      round(H5.exp.news.mod6.trends.tab[,c(5,6)],2))
H5.exp.news.mod6.trends.tab

write.csv2(H5.exp.news.mod6.trends.tab,"H5.exp.news.mod6.trends.tab.csv")



#contrast between anti-immigration and pro-environment
(H5.exp.news.contrast<-data.frame(pairs(H5.exp.news.mod6.trends, exclude = c(2:9),reverse=T)))
#contrast for all groups against mean of other groups
contrast(H5.exp.news.mod6.trends, "del.eff", by = NULL,adjust=c("holm"))
#contrast for three voting groups
(H5.exp.news.more.contrasts<-data.frame(pairs(H5.exp.news.mod6.trends, 
         exclude=c(2:8), by = NULL,adjust=c("holm"),reverse=T)))



```

