---
title: "Exploratory non-linearity"
output: 
  html_document: 
    keep_md: yes
    toc: yes
    toc_depth: 5
---


# Exploratory analysis: Is the association linear?

## Data re-preparations

```{r}

##Remove two of the smallest voting groups

Rdat<-dat %>%
  filter(all.parties.lvl2!="Invalid vote" & all.parties.lvl2!="No answer")
```

## Model 1 (H1 selected model without random effect correlations)

```{r}
EX9RR.NCmod1<-lmer(refugees~
                     (environ.lvl1||voting.group)+
                     (environ.lvl1||cntry)+
                     age+gender+educ+resid+occup+
                     environ.lvl1,data=Rdat,REML=F,
                   control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e8)))

isSingular(EX9RR.NCmod1)
(VC.EX9RR.NCmod1<-getVC(EX9RR.NCmod1))
getFE(EX9RR.NCmod1)
```

\newpage

## Model 2 (add squared fixed effect)

```{r}
EX9RR.NCmod2<-lmer(refugees~(environ.lvl1||voting.group)+
                     (environ.lvl1||cntry)+
                     age+gender+educ+resid+occup+
                     environ.lvl1+
                     I(environ.lvl1^2),data=Rdat,REML=F,
                   control=lmerControl(optimizer="bobyqa",
                                       optCtrl=list(maxfun=2e8)))


anova(EX9RR.NCmod1,EX9RR.NCmod2)
isSingular(EX9RR.NCmod2)
(VC.EX9RR.NCmod2<-getVC(EX9RR.NCmod2))
getFE(EX9RR.NCmod2)
```

\newpage

## Model 3 (add fixed cubic effect)

```{r}
EX9RR.NCmod3<-lmer(refugees~(environ.lvl1||voting.group)+
                     (environ.lvl1||cntry)+
                     age+gender+educ+resid+occup+
                     environ.lvl1+
            I(environ.lvl1^2)+
              I(environ.lvl1^3),
            data=Rdat,REML=F,
                   control=lmerControl(optimizer="bobyqa",
                                       optCtrl=list(maxfun=2e8)))

isSingular(EX9RR.NCmod3)
#test for cubic term only
anova(EX9RR.NCmod2,EX9RR.NCmod3)
#test for both non-linear terms
anova(EX9RR.NCmod1,EX9RR.NCmod3)

(VC.EX9RR.NCmod3<-getVC(EX9RR.NCmod3))
getFE(EX9RR.NCmod3)
```

## Model 4 (add random effects for non-linear terms by country)

```{r}
EX9RR.NCmod4<-lmer(refugees~(environ.lvl1||voting.group)+
                (environ.lvl1+
                   I(environ.lvl1^2)+
                   I(environ.lvl1^3)||cntry)+
                     age+gender+educ+resid+occup+
                     environ.lvl1+
                  I(environ.lvl1^2)+
                  I(environ.lvl1^3),
                data=Rdat,REML=F,
                   control=lmerControl(optimizer="bobyqa",
                                       optCtrl=list(maxfun=2e8)))

isSingular(EX9RR.NCmod4)
anova(EX9RR.NCmod3,EX9RR.NCmod4)
(VC.EX9RR.NCmod4<-getVC(EX9RR.NCmod4))
getFE(EX9RR.NCmod4)
```
## Model 5 (add random effects for non-linear terms by voting group)

```{r}
EX9RR.NCmod6<-lmer(refugees~
                     (environ.lvl1+
                        I(environ.lvl1^2)+
                        I(environ.lvl1^3)||voting.group)+
                     (environ.lvl1+
                        I(environ.lvl1^2)+
                        I(environ.lvl1^3)||cntry)+
                     age+gender+educ+resid+occup+
                     environ.lvl1+
                     I(environ.lvl1^2)+
                     I(environ.lvl1^3),data=Rdat,REML=F,
                   control=lmerControl(optimizer="bobyqa",
                                       optCtrl=list(maxfun=2e8)))

isSingular(EX9RR.NCmod6)
#test against fixed effects only
anova(EX9RR.NCmod3,EX9RR.NCmod6)
#test against random effects by country
anova(EX9RR.NCmod4,EX9RR.NCmod6)
(VC.EX9RR.NCmod6<-getVC(EX9RR.NCmod6))
getFE(EX9RR.NCmod6)
```

\newpage

### Marginal effects for non-linearity at level-1 

```{r message=FALSE, warning=FALSE}
EX9RR.NCmod6.trends<-
  emtrends(EX9RR.NCmod6,
           specs = c("environ.lvl1"),
           var=c("environ.lvl1"),
           at=list(environ.lvl1=
                     c(mean(Rdat$environ.lvl1)-1*sd(Rdat$environ.lvl1),
                       mean(Rdat$environ.lvl1)-0*sd(Rdat$environ.lvl1),
                       mean(Rdat$environ.lvl1)+1*sd(Rdat$environ.lvl1))))

(EX9RR.NCmod6.trends.tab<-data.frame(EX9RR.NCmod6.trends))


EX9RR.NCmod6.trends.tab$p<-
  2*(1-pnorm(abs(EX9RR.NCmod6.trends.tab$environ.lvl1.trend/
                   EX9RR.NCmod6.trends.tab$SE)))
EX9RR.NCmod6.trends.tab$adj.p<-
  p.adjust(EX9RR.NCmod6.trends.tab$p,method="holm")

EX9RR.NCmod6.trends.tab<-
  cbind(group=EX9RR.NCmod6.trends.tab[,1],
        round(EX9RR.NCmod6.trends.tab[,c(2,3)],2),
        round(EX9RR.NCmod6.trends.tab[,c(7,8)],4),
        round(EX9RR.NCmod6.trends.tab[,c(5,6)],2))
EX9RR.NCmod6.trends.tab

write.csv2(EX9RR.NCmod6.trends.tab,"EX9RR.NCmod6.trends.tab.csv")
```

\newpage

## Model 6 (include voting group main effects at level-2)

```{r}
EX9RR.NCmod7<-lmer(refugees~
                     (environ.lvl1+
                        I(environ.lvl1^2)+
                        I(environ.lvl1^3)||voting.group)+
                     (environ.lvl1+
                        I(environ.lvl1^2)+
                        I(environ.lvl1^3)||cntry)+
                     age+gender+educ+resid+occup+
                     environ.lvl1+
                     I(environ.lvl1^2)+
                     I(environ.lvl1^3)+
                     all.parties.lvl2,data=Rdat,REML=F,
                   control=lmerControl(optimizer="bobyqa",
                                       optCtrl=list(maxfun=2e8)))
isSingular(EX9RR.NCmod7)
anova(EX9RR.NCmod6,EX9RR.NCmod7)
getFE(EX9RR.NCmod7)
getVC(EX9RR.NCmod7)
```

## Model 7 (voting group interactions with linear term)

```{r}

EX9RR.NCmod8<-lmer(refugees~
                     (environ.lvl1+
                        I(environ.lvl1^2)+
                        I(environ.lvl1^3)||voting.group)+
                     (environ.lvl1+
                        I(environ.lvl1^2)+
                        I(environ.lvl1^3)||cntry)+
                     age+gender+educ+resid+occup+
                     environ.lvl1+
                     I(environ.lvl1^2)+
                     I(environ.lvl1^3)+
                     all.parties.lvl2+
                     all.parties.lvl2:environ.lvl1,
                   data=Rdat,REML=F,
                   control=lmerControl(optimizer="bobyqa",
                                       optCtrl=list(maxfun=2e8)))
isSingular(EX9RR.NCmod8)
anova(EX9RR.NCmod7,EX9RR.NCmod8)
getFE(EX9RR.NCmod8)
getVC(EX9RR.NCmod8)
```

\newpage

## Model 8 (interactions with non-linear terms)

```{r}
EX9RR.NCmod10<-lmer(refugees~
                      (environ.lvl1+
                         I(environ.lvl1^2)+
                         I(environ.lvl1^3)||voting.group)+
                      (environ.lvl1+
                         I(environ.lvl1^2)+
                         I(environ.lvl1^3)||cntry)+
                      age+gender+educ+resid+occup+
                      environ.lvl1+
                      I(environ.lvl1^2)+
                      I(environ.lvl1^3)+
                      all.parties.lvl2+
                      all.parties.lvl2:environ.lvl1+
                      all.parties.lvl2:I(environ.lvl1^2)+
                      all.parties.lvl2:I(environ.lvl1^3),
                    data=Rdat,REML=F,
                    control=lmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e8)))

#test against linear interaction by voting groups
anova(EX9RR.NCmod8,EX9RR.NCmod10)

getFE(EX9RR.NCmod10)
getVC(EX9RR.NCmod10)
```

\newpage

### Marginal trends for each voting group

#### Linear coefficients

```{r message=FALSE, warning=FALSE}

EX9RR.NCmod10.linear<-
  emtrends(EX9RR.NCmod10,specs = c("all.parties.lvl2"),var=c("environ.lvl1"))
(EX9RR.NCmod10.linear.tab<-data.frame(EX9RR.NCmod10.linear))


EX9RR.NCmod10.linear.tab$p<-
  2*(1-pnorm(abs(EX9RR.NCmod10.linear.tab$environ.lvl1.trend/
                   EX9RR.NCmod10.linear.tab$SE)))
EX9RR.NCmod10.linear.tab$adj.p<-
  p.adjust(EX9RR.NCmod10.linear.tab$p,method="holm")

EX9RR.NCmod10.linear.tab<-
  cbind(group=EX9RR.NCmod10.linear.tab[,1],
        round(EX9RR.NCmod10.linear.tab[,c(2,3)],2),
        round(EX9RR.NCmod10.linear.tab[,c(7,8)],4),
        round(EX9RR.NCmod10.linear.tab[,c(5,6)],2))
EX9RR.NCmod10.linear.tab

write.csv2(EX9RR.NCmod10.linear.tab,"EX9RR.NCmod10.linear.tab.csv")

#contrast for three voting groups
(EX9RR.more.contrasts<-data.frame(pairs(EX9RR.NCmod10.linear, 
                                        exclude=c(2:6), by = NULL,adjust=c("none"),reverse=T)))


```

### Quadratic and cubic coefficients

* Non-linear variables must be manually recoded to new variables for extracting marginal effects

```{r}
#recoding
Rdat$environ.lvl1.sq<-Rdat$environ.lvl1^2
Rdat$environ.lvl1.cu<-Rdat$environ.lvl1^3

#refitting the model
EX9RR.NCmod10.ma<-lmer(refugees~
                         (environ.lvl1+
                            environ.lvl1.sq+
                            environ.lvl1.cu||voting.group)+
                         (environ.lvl1+
                            environ.lvl1.sq+
                            environ.lvl1.cu||cntry)+
                         age+gender+educ+resid+occup+
                         environ.lvl1+
                         environ.lvl1.sq+
                         environ.lvl1.cu+
                         all.parties.lvl2+
                         all.parties.lvl2:environ.lvl1+
                         all.parties.lvl2:environ.lvl1.sq+
                       all.parties.lvl2:environ.lvl1.cu,
                       data=Rdat,REML=F,
                       control=lmerControl(optimizer="bobyqa",
                                           optCtrl=list(maxfun=2e8)))
isSingular(EX9RR.NCmod10.ma)
#check if identical
anova(EX9RR.NCmod10,EX9RR.NCmod10.ma)
#
anova(EX9RR.NCmod8,EX9RR.NCmod10.ma)

getFE(EX9RR.NCmod10.ma)
getVC(EX9RR.NCmod10.ma)
```

#### Marginal quadratic effects

```{r message=FALSE, warning=FALSE}

EX9RR.NCmod10.quadratic<-
  emtrends(EX9RR.NCmod10.ma,specs = c("all.parties.lvl2"),var=c("environ.lvl1.sq"))
(EX9RR.NCmod10.quadratic.tab<-data.frame(EX9RR.NCmod10.quadratic))


EX9RR.NCmod10.quadratic.tab$p<-
  2*(1-pnorm(abs(EX9RR.NCmod10.quadratic.tab$environ.lvl1.sq.trend/
                   EX9RR.NCmod10.quadratic.tab$SE)))
EX9RR.NCmod10.quadratic.tab$adj.p<-
  p.adjust(EX9RR.NCmod10.quadratic.tab$p,method="holm")

EX9RR.NCmod10.quadratic.tab<-
  cbind(group=EX9RR.NCmod10.quadratic.tab[,1],
        round(EX9RR.NCmod10.quadratic.tab[,c(2,3)],2),
        round(EX9RR.NCmod10.quadratic.tab[,c(7,8)],4),
        round(EX9RR.NCmod10.quadratic.tab[,c(5,6)],2))
EX9RR.NCmod10.quadratic.tab

write.csv2(EX9RR.NCmod10.quadratic.tab,"EX9RR.NCmod10.quadratic.tab.csv")

#contrast for three voting groups
(EX9RR.more.contrasts<-data.frame(pairs(EX9RR.NCmod10.quadratic, 
                                        exclude=c(2:6), by = NULL,adjust=c("none"),reverse=T)))
(EX9RR.more.contrasts<-data.frame(pairs(EX9RR.NCmod10.quadratic, "del.eff",
                                        exclude=c(2:6), by = NULL,adjust=c("none"),reverse=T)))
```

\newpage

#### Marginal cubic effects

```{r message=FALSE, warning=FALSE}

EX9RR.NCmod10.cubic<-
  emtrends(EX9RR.NCmod10.ma,specs = c("all.parties.lvl2"),var=c("environ.lvl1.cu"))
(EX9RR.NCmod10.cubic.tab<-data.frame(EX9RR.NCmod10.cubic))


EX9RR.NCmod10.cubic.tab$p<-
  2*(1-pnorm(abs(EX9RR.NCmod10.cubic.tab$environ.lvl1.cu.trend/
                   EX9RR.NCmod10.cubic.tab$SE)))
EX9RR.NCmod10.cubic.tab$adj.p<-
  p.adjust(EX9RR.NCmod10.cubic.tab$p,method="holm")

EX9RR.NCmod10.cubic.tab<-
  cbind(group=EX9RR.NCmod10.cubic.tab[,1],
        round(EX9RR.NCmod10.cubic.tab[,c(2,3)],2),
        round(EX9RR.NCmod10.cubic.tab[,c(7,8)],4),
        round(EX9RR.NCmod10.cubic.tab[,c(5,6)],2))
EX9RR.NCmod10.cubic.tab

write.csv2(EX9RR.NCmod10.cubic.tab,"EX9RR.NCmod10.cubic.tab.csv")

#contrast for three voting groups
(EX9RR.more.contrasts<-data.frame(pairs(EX9RR.NCmod10.cubic, 
                                        exclude=c(2:6), by = NULL,adjust=c("none"),reverse=T)))
(EX9RR.more.contrasts<-data.frame(pairs(EX9RR.NCmod10.cubic, "del.eff",
                                        exclude=c(2:6), by = NULL,adjust=c("none"),reverse=T)))

```

\newpage

### Simple slopes for each voting group at -1SD, mean, and +1SD (grand mean used to enhance comparisons between voting groups)

* Because the location of points on grand-mean vary for group-means, the locations of grand mean points for each voting group are calculated separately

#### This script would produce the slopes for group means (not interpreted in the main text)

```{r message=FALSE, warning=FALSE}
EX9RR.NCmod10.slopes<-emtrends(EX9RR.NCmod10,specs = c("environ.lvl1","all.parties.lvl2"),var="environ.lvl1",
                               at=list(environ.lvl1=c(mean(dat$environ.lvl1)-sd(dat$environ.lvl1),
                                                      mean(dat$environ.lvl1)-0*sd(dat$environ.lvl1),
                                                      mean(dat$environ.lvl1)+sd(dat$environ.lvl1))))
(EX9RR.NCmod10.slopes.tab<-data.frame(EX9RR.NCmod10.slopes))


EX9RR.NCmod10.slopes.tab$p<-
  2*(1-pnorm(abs(EX9RR.NCmod10.slopes.tab$environ.lvl1.trend/
                   EX9RR.NCmod10.slopes.tab$SE)))
EX9RR.NCmod10.slopes.tab$adj.p<-
  p.adjust(EX9RR.NCmod10.slopes.tab$p,method="holm")

EX9RR.NCmod10.slopes.tab<-
  cbind(env_point=EX9RR.NCmod10.slopes.tab[,1],group=EX9RR.NCmod10.slopes.tab[,2],
        round(EX9RR.NCmod10.slopes.tab[,c(3,4)],2),
        round(EX9RR.NCmod10.slopes.tab[,c(8,9)],4),
        round(EX9RR.NCmod10.slopes.tab[,c(6,7)],2))
EX9RR.NCmod10.slopes.tab


write.csv2(EX9RR.NCmod10.slopes.tab,"EX9RR.NCmod10.slopes.tab.csv")
```

\newpage

#### Locating the grand-mean points for group-means

```{r}

all.parties.lvl2.means<-Rdat %>%
  group_by(all.parties.lvl2) %>%
  summarize(env.lvl2.mean=mean(environ.cntrymc),
            env.lvl1.sd=sd(environ.lvl1),
            ref.lvl2.mean=mean(refugees.cntrymc),
            ref.lvl1.sd=sd(refugees.lvl1),
            n=n())

all.parties.lvl2.means<-data.frame(all.parties.lvl2.means)

all.parties.lvl2.means$env.num<-
  (all.parties.lvl2.means$n-1)*
  (all.parties.lvl2.means$env.lvl1.sd^2)

all.parties.lvl2.means$env.pooled.sd<-
  sqrt(sum(all.parties.lvl2.means$env.num)/
         (sum(all.parties.lvl2.means$n)-
            nrow(all.parties.lvl2.means)))

all.parties.lvl2.means$ref.num<-
  (all.parties.lvl2.means$n-1)*
  (all.parties.lvl2.means$ref.lvl1.sd^2)

all.parties.lvl2.means$ref.pooled.sd<-
  sqrt(sum(all.parties.lvl2.means$ref.num)/
         (sum(all.parties.lvl2.means$n)-
            nrow(all.parties.lvl2.means)))
all.parties.lvl2.means

```

##### group.1: Anti-immigration party

```{r message=FALSE, warning=FALSE}

x.points.group.1<-
  c(mean(dat$environ.lvl1)-1*all.parties.lvl2.means$env.pooled.sd[1]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[1]),
    mean(dat$environ.lvl1)-0*all.parties.lvl2.means$env.pooled.sd[1]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[1]),
    mean(dat$environ.lvl1)+1*all.parties.lvl2.means$env.pooled.sd[1]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[1]))

x.points.group.1

EX9RR.NCmod10.slopes.group.1<-emtrends(EX9RR.NCmod10,specs = c("environ.lvl1","all.parties.lvl2"),var="environ.lvl1",
                                       at=list(environ.lvl1=c(x.points.group.1)))
EX9RR.NCmod10.slopes.group.1.tab<-data.frame(EX9RR.NCmod10.slopes.group.1)
(EX9RR.NCmod10.slopes.group.1.tab<-
    EX9RR.NCmod10.slopes.group.1.tab[EX9RR.NCmod10.slopes.group.1.tab$all.parties.lvl2=="Anti-immigration party",])

EX9RR.NCmod10.slopes.group.1.tab$p<-
  2*(1-pnorm(abs(EX9RR.NCmod10.slopes.group.1.tab$environ.lvl1.trend/
                   EX9RR.NCmod10.slopes.group.1.tab$SE)))
EX9RR.NCmod10.slopes.group.1.tab$adj.p<-
  p.adjust(EX9RR.NCmod10.slopes.group.1.tab$p,method="holm")

EX9RR.NCmod10.slopes.group.1.tab<-
  cbind(env_point=EX9RR.NCmod10.slopes.group.1.tab[,1],group=EX9RR.NCmod10.slopes.group.1.tab[,2],
        round(EX9RR.NCmod10.slopes.group.1.tab[,c(3,4)],2),
        round(EX9RR.NCmod10.slopes.group.1.tab[,c(8,9)],4),
        round(EX9RR.NCmod10.slopes.group.1.tab[,c(6,7)],2))
EX9RR.NCmod10.slopes.group.1.tab
```

\newpage

##### group.2: Did not vote -group

```{r message=FALSE, warning=FALSE}

x.points.group.2<-
  c(mean(dat$environ.lvl1)-1*all.parties.lvl2.means$env.pooled.sd[2]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[2]),
    mean(dat$environ.lvl1)-0*all.parties.lvl2.means$env.pooled.sd[2]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[2]),
    mean(dat$environ.lvl1)+1*all.parties.lvl2.means$env.pooled.sd[2]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[2]))

x.points.group.2

EX9RR.NCmod10.slopes.group.2<-emtrends(EX9RR.NCmod10,specs = c("environ.lvl1","all.parties.lvl2"),var="environ.lvl1",
                                       at=list(environ.lvl1=c(x.points.group.2)))
EX9RR.NCmod10.slopes.group.2.tab<-data.frame(EX9RR.NCmod10.slopes.group.2)
(EX9RR.NCmod10.slopes.group.2.tab<-
    EX9RR.NCmod10.slopes.group.2.tab[EX9RR.NCmod10.slopes.group.2.tab$all.parties.lvl2=="Did not vote",])

EX9RR.NCmod10.slopes.group.2.tab$p<-
  2*(1-pnorm(abs(EX9RR.NCmod10.slopes.group.2.tab$environ.lvl1.trend/
                   EX9RR.NCmod10.slopes.group.2.tab$SE)))
EX9RR.NCmod10.slopes.group.2.tab$adj.p<-
  p.adjust(EX9RR.NCmod10.slopes.group.2.tab$p,method="holm")

EX9RR.NCmod10.slopes.group.2.tab<-
  cbind(env_point=EX9RR.NCmod10.slopes.group.2.tab[,1],group=EX9RR.NCmod10.slopes.group.2.tab[,2],
        round(EX9RR.NCmod10.slopes.group.2.tab[,c(3,4)],2),
        round(EX9RR.NCmod10.slopes.group.2.tab[,c(8,9)],4),
        round(EX9RR.NCmod10.slopes.group.2.tab[,c(6,7)],2))
EX9RR.NCmod10.slopes.group.2.tab
```

\newpage

##### group.3: Don't know -group

```{r message=FALSE, warning=FALSE}

x.points.group.3<-
  c(mean(dat$environ.lvl1)-1*all.parties.lvl2.means$env.pooled.sd[3]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[3]),
    mean(dat$environ.lvl1)-0*all.parties.lvl2.means$env.pooled.sd[3]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[3]),
    mean(dat$environ.lvl1)+1*all.parties.lvl2.means$env.pooled.sd[3]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[3]))

x.points.group.3

EX9RR.NCmod10.slopes.group.3<-emtrends(EX9RR.NCmod10,specs = c("environ.lvl1","all.parties.lvl2"),var="environ.lvl1",
                                       at=list(environ.lvl1=c(x.points.group.3)))
EX9RR.NCmod10.slopes.group.3.tab<-data.frame(EX9RR.NCmod10.slopes.group.3)
(EX9RR.NCmod10.slopes.group.3.tab<-
    EX9RR.NCmod10.slopes.group.3.tab[EX9RR.NCmod10.slopes.group.3.tab$all.parties.lvl2=="Don't know",])

EX9RR.NCmod10.slopes.group.3.tab$p<-
  2*(1-pnorm(abs(EX9RR.NCmod10.slopes.group.3.tab$environ.lvl1.trend/
                   EX9RR.NCmod10.slopes.group.3.tab$SE)))
EX9RR.NCmod10.slopes.group.3.tab$adj.p<-
  p.adjust(EX9RR.NCmod10.slopes.group.3.tab$p,method="holm")

EX9RR.NCmod10.slopes.group.3.tab<-
  cbind(env_point=EX9RR.NCmod10.slopes.group.3.tab[,1],group=EX9RR.NCmod10.slopes.group.3.tab[,2],
        round(EX9RR.NCmod10.slopes.group.3.tab[,c(3,4)],2),
        round(EX9RR.NCmod10.slopes.group.3.tab[,c(8,9)],4),
        round(EX9RR.NCmod10.slopes.group.3.tab[,c(6,7)],2))
EX9RR.NCmod10.slopes.group.3.tab
```

\newpage

##### group.4: Not eligible Age -group

```{r message=FALSE, warning=FALSE}

x.points.group.4<-
  c(mean(dat$environ.lvl1)-1*all.parties.lvl2.means$env.pooled.sd[4]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[4]),
    mean(dat$environ.lvl1)-0*all.parties.lvl2.means$env.pooled.sd[4]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[4]),
    mean(dat$environ.lvl1)+1*all.parties.lvl2.means$env.pooled.sd[4]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[4]))

x.points.group.4

EX9RR.NCmod10.slopes.group.4<-emtrends(EX9RR.NCmod10,specs = c("environ.lvl1","all.parties.lvl2"),var="environ.lvl1",
                                       at=list(environ.lvl1=c(x.points.group.4)))
EX9RR.NCmod10.slopes.group.4.tab<-data.frame(EX9RR.NCmod10.slopes.group.4)
(EX9RR.NCmod10.slopes.group.4.tab<-
    EX9RR.NCmod10.slopes.group.4.tab[EX9RR.NCmod10.slopes.group.4.tab$all.parties.lvl2=="NE age",])

EX9RR.NCmod10.slopes.group.4.tab$p<-
  2*(1-pnorm(abs(EX9RR.NCmod10.slopes.group.4.tab$environ.lvl1.trend/
                   EX9RR.NCmod10.slopes.group.4.tab$SE)))
EX9RR.NCmod10.slopes.group.4.tab$adj.p<-
  p.adjust(EX9RR.NCmod10.slopes.group.4.tab$p,method="holm")

EX9RR.NCmod10.slopes.group.4.tab<-
  cbind(env_point=EX9RR.NCmod10.slopes.group.4.tab[,1],group=EX9RR.NCmod10.slopes.group.4.tab[,2],
        round(EX9RR.NCmod10.slopes.group.4.tab[,c(3,4)],2),
        round(EX9RR.NCmod10.slopes.group.4.tab[,c(8,9)],4),
        round(EX9RR.NCmod10.slopes.group.4.tab[,c(6,7)],2))
EX9RR.NCmod10.slopes.group.4.tab
```

\newpage

##### group.5: Not eligible citizenship -group

```{r message=FALSE, warning=FALSE}

x.points.group.5<-
  c(mean(dat$environ.lvl1)-1*all.parties.lvl2.means$env.pooled.sd[5]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[5]),
    mean(dat$environ.lvl1)-0*all.parties.lvl2.means$env.pooled.sd[5]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[5]),
    mean(dat$environ.lvl1)+1*all.parties.lvl2.means$env.pooled.sd[5]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[5]))

x.points.group.5

EX9RR.NCmod10.slopes.group.5<-emtrends(EX9RR.NCmod10,specs = c("environ.lvl1","all.parties.lvl2"),var="environ.lvl1",
                                       at=list(environ.lvl1=c(x.points.group.5)))
EX9RR.NCmod10.slopes.group.5.tab<-data.frame(EX9RR.NCmod10.slopes.group.5)
(EX9RR.NCmod10.slopes.group.5.tab<-
    EX9RR.NCmod10.slopes.group.5.tab[EX9RR.NCmod10.slopes.group.5.tab$all.parties.lvl2=="NE citizen",])

EX9RR.NCmod10.slopes.group.5.tab$p<-
  2*(1-pnorm(abs(EX9RR.NCmod10.slopes.group.5.tab$environ.lvl1.trend/
                   EX9RR.NCmod10.slopes.group.5.tab$SE)))
EX9RR.NCmod10.slopes.group.5.tab$adj.p<-
  p.adjust(EX9RR.NCmod10.slopes.group.5.tab$p,method="holm")

EX9RR.NCmod10.slopes.group.5.tab<-
  cbind(env_point=EX9RR.NCmod10.slopes.group.5.tab[,1],group=EX9RR.NCmod10.slopes.group.5.tab[,2],
        round(EX9RR.NCmod10.slopes.group.5.tab[,c(3,4)],2),
        round(EX9RR.NCmod10.slopes.group.5.tab[,c(8,9)],4),
        round(EX9RR.NCmod10.slopes.group.5.tab[,c(6,7)],2))
EX9RR.NCmod10.slopes.group.5.tab
```

\newpage

##### group.6: Not eligible Other -group

```{r message=FALSE, warning=FALSE}

x.points.group.6<-
  c(mean(dat$environ.lvl1)-1*all.parties.lvl2.means$env.pooled.sd[6]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[6]),
    mean(dat$environ.lvl1)-0*all.parties.lvl2.means$env.pooled.sd[6]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[6]),
    mean(dat$environ.lvl1)+1*all.parties.lvl2.means$env.pooled.sd[6]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[6]))

x.points.group.6

EX9RR.NCmod10.slopes.group.6<-emtrends(EX9RR.NCmod10,specs = c("environ.lvl1","all.parties.lvl2"),var="environ.lvl1",
                                       at=list(environ.lvl1=c(x.points.group.6)))
EX9RR.NCmod10.slopes.group.6.tab<-data.frame(EX9RR.NCmod10.slopes.group.6)
(EX9RR.NCmod10.slopes.group.6.tab<-
    EX9RR.NCmod10.slopes.group.6.tab[EX9RR.NCmod10.slopes.group.6.tab$all.parties.lvl2=="NE other",])

EX9RR.NCmod10.slopes.group.6.tab$p<-
  2*(1-pnorm(abs(EX9RR.NCmod10.slopes.group.6.tab$environ.lvl1.trend/
                   EX9RR.NCmod10.slopes.group.6.tab$SE)))
EX9RR.NCmod10.slopes.group.6.tab$adj.p<-
  p.adjust(EX9RR.NCmod10.slopes.group.6.tab$p,method="holm")

EX9RR.NCmod10.slopes.group.6.tab<-
  cbind(env_point=EX9RR.NCmod10.slopes.group.6.tab[,1],group=EX9RR.NCmod10.slopes.group.6.tab[,2],
        round(EX9RR.NCmod10.slopes.group.6.tab[,c(3,4)],2),
        round(EX9RR.NCmod10.slopes.group.6.tab[,c(8,9)],4),
        round(EX9RR.NCmod10.slopes.group.6.tab[,c(6,7)],2))
EX9RR.NCmod10.slopes.group.6.tab
```

##### group.7: Other party -group

```{r message=FALSE, warning=FALSE}

x.points.group.7<-
  c(mean(dat$environ.lvl1)-1*all.parties.lvl2.means$env.pooled.sd[7]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[7]),
    mean(dat$environ.lvl1)-0*all.parties.lvl2.means$env.pooled.sd[7]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[7]),
    mean(dat$environ.lvl1)+1*all.parties.lvl2.means$env.pooled.sd[7]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[7]))

x.points.group.7

EX9RR.NCmod10.slopes.group.7<-emtrends(EX9RR.NCmod10,specs = c("environ.lvl1","all.parties.lvl2"),var="environ.lvl1",
                                       at=list(environ.lvl1=c(x.points.group.7)))
EX9RR.NCmod10.slopes.group.7.tab<-data.frame(EX9RR.NCmod10.slopes.group.7)
(EX9RR.NCmod10.slopes.group.7.tab<-
    EX9RR.NCmod10.slopes.group.7.tab[EX9RR.NCmod10.slopes.group.7.tab$all.parties.lvl2=="Other party",])

EX9RR.NCmod10.slopes.group.7.tab$p<-
  2*(1-pnorm(abs(EX9RR.NCmod10.slopes.group.7.tab$environ.lvl1.trend/
                   EX9RR.NCmod10.slopes.group.7.tab$SE)))
EX9RR.NCmod10.slopes.group.7.tab$adj.p<-
  p.adjust(EX9RR.NCmod10.slopes.group.7.tab$p,method="holm")

EX9RR.NCmod10.slopes.group.7.tab<-
  cbind(env_point=EX9RR.NCmod10.slopes.group.7.tab[,1],group=EX9RR.NCmod10.slopes.group.7.tab[,2],
        round(EX9RR.NCmod10.slopes.group.7.tab[,c(3,4)],2),
        round(EX9RR.NCmod10.slopes.group.7.tab[,c(8,9)],4),
        round(EX9RR.NCmod10.slopes.group.7.tab[,c(6,7)],2))
EX9RR.NCmod10.slopes.group.7.tab
```

\newpage

##### group.8: Pro-environment party -group

```{r message=FALSE, warning=FALSE}

x.points.group.8<-
  c(mean(dat$environ.lvl1)-1*all.parties.lvl2.means$env.pooled.sd[8]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[8]),
    mean(dat$environ.lvl1)-0*all.parties.lvl2.means$env.pooled.sd[8]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[8]),
    mean(dat$environ.lvl1)+1*all.parties.lvl2.means$env.pooled.sd[8]+
      (mean(dat$environ.lvl1)-all.parties.lvl2.means$env.lvl2.mean[8]))

x.points.group.8

EX9RR.NCmod10.slopes.group.8<-emtrends(EX9RR.NCmod10,specs = c("environ.lvl1","all.parties.lvl2"),var="environ.lvl1",
                                       at=list(environ.lvl1=c(x.points.group.8)))
EX9RR.NCmod10.slopes.group.8.tab<-data.frame(EX9RR.NCmod10.slopes.group.8)
(EX9RR.NCmod10.slopes.group.8.tab<-
    EX9RR.NCmod10.slopes.group.8.tab[EX9RR.NCmod10.slopes.group.8.tab$all.parties.lvl2=="Pro-environment party",])

EX9RR.NCmod10.slopes.group.8.tab$p<-
  2*(1-pnorm(abs(EX9RR.NCmod10.slopes.group.8.tab$environ.lvl1.trend/
                   EX9RR.NCmod10.slopes.group.8.tab$SE)))
EX9RR.NCmod10.slopes.group.8.tab$adj.p<-
  p.adjust(EX9RR.NCmod10.slopes.group.8.tab$p,method="holm")

EX9RR.NCmod10.slopes.group.8.tab<-
  cbind(env_point=EX9RR.NCmod10.slopes.group.8.tab[,1],group=EX9RR.NCmod10.slopes.group.8.tab[,2],
        round(EX9RR.NCmod10.slopes.group.8.tab[,c(3,4)],2),
        round(EX9RR.NCmod10.slopes.group.8.tab[,c(8,9)],4),
        round(EX9RR.NCmod10.slopes.group.8.tab[,c(6,7)],2))
EX9RR.NCmod10.slopes.group.8.tab
```

### Plotting the effects

#### Make a separate long format data file for the focal groups + young non-voters

```{r}
EX9.plot.dat<-Rdat %>%
  filter(anti.imm.party.dummy==1 |
           pro.env.party.dummy==1 |
           other.party.dummy==1 |
           not.eligible.age.dummy==1)

#rename the groups
EX9.plot.dat$party<-ifelse(EX9.plot.dat$anti.imm.party.dummy==1,"Anti-immigration voters",
                           ifelse(EX9.plot.dat$pro.env.party.dummy==1,"Pro-environment voters",
                                  ifelse(EX9.plot.dat$other.party.dummy==1,"Other party voters",
                                         ifelse(EX9.plot.dat$not.eligible.age.dummy==1,"Not eligible to vote: Age",NA))))
```

\newpage

#### Plot the curves

```{r}

non.lin.plot<-
  ggplot(data=EX9.plot.dat,aes(x=environ.gmc,y=refugees))+
  geom_smooth(method="lm",
              se=T,
              formula=y ~ poly(x, 3),
              size=2,color="black")+
  geom_smooth(method="lm",
              se=F,
              formula=y ~ x,
              size=1,
              color="red",linetype="dashed")+
  xlab("Attitudes towards the Environment")+
  ylab("Attitudes towards Refugees")+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"),
        plot.title = element_text(size = 16, face = "bold"),
        legend.title=element_text(size=14), 
        legend.text=element_text(size=14),
        panel.background = element_blank(),
        panel.grid.major = element_line(size = 0.25,
                                        linetype = 'dotted',
                                        colour = "black"))+
  coord_cartesian(xlim=c(-2,2))+
  facet_wrap(~party,ncol=2)+
  theme(strip.text.x = element_text(size = 12,face="bold"),
        strip.background = element_blank())

non.lin.plot

ggsave(plot = non.lin.plot,
       filename="non.lin.plot.png",device = "png",
       units = "cm",width=12,height=18,dpi = 600)

```

