---
title: "Hypothesis 4"
output: html_document
---


# Hypothesis 4: The strength of the association between environment and refugee attitudes is stronger among those who voted for pro-environment or anti-immigration parties in the previous national elections

### Model 1: without interactions (only main effects, combine H1 and H2 final models in terms of predictors)

```{r}
H4.mod1<-lmer(refugees~(environ.lvl1|voting.group)+
                (environ.lvl1|cntry)+
                age+gender+educ+resid+occup+
                environ.lvl1+
                all.parties.lvl2,data=dat,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))


isSingular(H4.mod1)
(FE.H4.mod1<-getFE(H4.mod1))
(VC.H4.mod1<-getVC(H4.mod1))


```

\newpage

### Model 2: Cross-level interaction between environmental attitudes and voting group categories

```{r}
H4.mod2<-lmer(refugees~(environ.lvl1|voting.group)+
                (environ.lvl1|cntry)+
                age+gender+educ+resid+occup+
                environ.lvl1+
                all.parties.lvl2+
                environ.lvl1:all.parties.lvl2,data=dat,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))
isSingular(H4.mod2)
anova(H4.mod1,H4.mod2)

(FE.H4.mod2<-getFE(H4.mod2))
(VC.H4.mod2<-getVC(H4.mod2))

```

\newpage

#### Marginal effect for pro-environment and anti-immigration voters

```{r message=FALSE}
H4.mod2.trends<-emtrends(H4.mod2,specs = c("all.parties.lvl2"),var=c("environ.lvl1"))
(H4.mod2.trends.tab<-data.frame(H4.mod2.trends))


H4.mod2.trends.tab$p<-
  2*(1-pnorm(abs(H4.mod2.trends.tab$environ.lvl1.trend/
                   H4.mod2.trends.tab$SE)))
H4.mod2.trends.tab$adj.p<-
  p.adjust(H4.mod2.trends.tab$p,method="holm")

H4.mod2.trends.tab<-
  cbind(group=H4.mod2.trends.tab[,1],
      round(H4.mod2.trends.tab[,c(2,3)],2),
      round(H4.mod2.trends.tab[,c(7,8)],4),
      round(H4.mod2.trends.tab[,c(5,6)],2))
H4.mod2.trends.tab

write.csv2(H4.mod2.trends.tab,"H4.mod2.trends.tab.csv")

#contrast between anti-immigration and pro-environment
(H4.contrast<-data.frame(pairs(H4.mod2.trends, exclude = c(2:9),reverse=T)))
#contrast for all groups against mean of other groups
contrast(H4.mod2.trends, "del.eff", by = NULL,adjust=c("none"))
#contrast for three voting groups
(H4.more.contrasts<-data.frame(pairs(H4.mod2.trends, 
         exclude=c(2:8), by = NULL,adjust=c("none"),reverse=T)))



```



\newpage
