---
title: "Preparations"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r message=FALSE, warning=FALSE}
library(lme4)
library(lmerTest)
library(dplyr)
library(psych)
library(emmeans)
library(ggplot2)
library(metafor)
library(merTools)
```

## Session information about the packages

```{r}
sessionInfo()
```

\newpage

## Custom functions

```{r}

#to extract fixed effects
getFE<-function(model){
  coefs<-data.frame(summary(model)$coefficients)
  coefs$lower<-coefs[,1]-qt(p=.975,df=coefs[,"df"])*coefs[,2]
  coefs$upper<-coefs[,1]+qt(p=.975,df=coefs[,"df"])*coefs[,2]
  coefs<-cbind.data.frame(round(coefs[,1:4],2),
                          p=round(coefs[,5],3),
                          LL=round(coefs$lower,2),
                          UL=round(coefs$upper,2))
  #row.names(coefs)<-substr(row.names(coefs),1,25)
  return(coefs)
}



#to extract random effects
getVC<-function(model){
  VC<-as.data.frame(VarCorr(model))
  VC<-cbind(VC[,c(1:3)],est_SD=VC[,5],est_SD2=VC[,4])
  return(VC)
}


#to extract model deviance
getDEV<-function(model){
  DEV<-unname(summary(model)$devcomp$cmp["dev"])
  return(DEV)
}


#partial correlation test
pcor.test <- function(x,y,z,use="mat",method="p",na.rm=T){
	# The partial correlation coefficient between x and y given z
	#
	# pcor.test is free and comes with ABSOLUTELY NO WARRANTY.
	#
	# x and y should be vectors
	#
	# z can be either a vector or a matrix
	#
	# use: There are two methods to calculate the partial correlation coefficient.
	#	 One is by using variance-covariance matrix ("mat") and the other is by using recursive formula ("rec").
	#	 Default is "mat".
	#
	# method: There are three ways to calculate the correlation coefficient, 
	#	    which are Pearson's ("p"), Spearman's ("s"), and Kendall's ("k") methods.
	# 	    The last two methods which are Spearman's and Kendall's coefficient are based on the non-parametric analysis.
	#	    Default is "p".
	#
	# na.rm: If na.rm is T, then all the missing samples are deleted from the whole dataset, which is (x,y,z).
	#        If not, the missing samples will be removed just when the correlation coefficient is calculated.
	#	   However, the number of samples for the p-value is the number of samples after removing 
	#	   all the missing samples from the whole dataset.
	#	   Default is "T".

	x <- c(x)
	y <- c(y)
	z <- as.data.frame(z)

	if(use == "mat"){
		p.use <- "Var-Cov matrix"
		pcor = pcor.mat(x,y,z,method=method,na.rm=na.rm)
	}else if(use == "rec"){
		p.use <- "Recursive formula"
		pcor = pcor.rec(x,y,z,method=method,na.rm=na.rm)
	}else{
		stop("\'use\' should be either \"rec\" or \"mat\"!\n")
	}

	# print the method
	if(gregexpr("p",method)[[1]][1] == 1){
		p.method <- "Pearson"
	}else if(gregexpr("s",method)[[1]][1] == 1){
		p.method <- "Spearman"
	}else if(gregexpr("k",method)[[1]][1] == 1){
		p.method <- "Kendall"
	}else{
		stop("\'method\' should be \"pearson\" or \"spearman\" or \"kendall\"!\n")
	}

	# sample number
	n <- dim(na.omit(data.frame(x,y,z)))[1]
	
	# given variables' number
	gn <- dim(z)[2]

	# p-value
	if(p.method == "Kendall"){
		statistic <- pcor/sqrt(2*(2*(n-gn)+5)/(9*(n-gn)*(n-1-gn)))
		p.value <- 2*pnorm(-abs(statistic))

	}else{
		statistic <- pcor*sqrt((n-2-gn)/(1-pcor^2))
  		p.value <- 2*pnorm(-abs(statistic))
	}

	data.frame(estimate=pcor,p.value=p.value,statistic=statistic,n=n,gn=gn,Method=p.method,Use=p.use)
}			


# By using var-cov matrix
pcor.mat <- function(x,y,z,method="p",na.rm=T){

	x <- c(x)
	y <- c(y)
	z <- as.data.frame(z)

	if(dim(z)[2] == 0){
		stop("There should be given data\n")
	}

	data <- data.frame(x,y,z)

	if(na.rm == T){
		data = na.omit(data)
	}

	xdata <- na.omit(data.frame(data[,c(1,2)]))
	Sxx <- cov(xdata,xdata,m=method)

	xzdata <- na.omit(data)
	xdata <- data.frame(xzdata[,c(1,2)])
	zdata <- data.frame(xzdata[,-c(1,2)])
	Sxz <- cov(xdata,zdata,m=method)

	zdata <- na.omit(data.frame(data[,-c(1,2)]))
	Szz <- cov(zdata,zdata,m=method)

	# is Szz positive definite?
	zz.ev <- eigen(Szz)$values
	if(min(zz.ev)[1]<0){
		stop("\'Szz\' is not positive definite!\n")
	}

	# partial correlation
	Sxx.z <- Sxx - Sxz %*% solve(Szz) %*% t(Sxz)
	
	rxx.z <- cov2cor(Sxx.z)[1,2]

	rxx.z
}

# By using recursive formula
pcor.rec <- function(x,y,z,method="p",na.rm=T){
	# 

	x <- c(x)
	y <- c(y)
	z <- as.data.frame(z)

	if(dim(z)[2] == 0){
		stop("There should be given data\n")
	}

	data <- data.frame(x,y,z)

	if(na.rm == T){
		data = na.omit(data)
	}

	# recursive formula
	if(dim(z)[2] == 1){
		tdata <- na.omit(data.frame(data[,1],data[,2]))
		rxy <- cor(tdata[,1],tdata[,2],m=method)

		tdata <- na.omit(data.frame(data[,1],data[,-c(1,2)]))
		rxz <- cor(tdata[,1],tdata[,2],m=method)

		tdata <- na.omit(data.frame(data[,2],data[,-c(1,2)]))
		ryz <- cor(tdata[,1],tdata[,2],m=method)

		rxy.z <- (rxy - rxz*ryz)/( sqrt(1-rxz^2)*sqrt(1-ryz^2) )
		
		return(rxy.z)
	}else{
		x <- c(data[,1])
		y <- c(data[,2])
		z0 <- c(data[,3])
		zc <- as.data.frame(data[,-c(1,2,3)])

		rxy.zc <- pcor.rec(x,y,zc,method=method,na.rm=na.rm)
		rxz0.zc <- pcor.rec(x,z0,zc,method=method,na.rm=na.rm)
		ryz0.zc <- pcor.rec(y,z0,zc,method=method,na.rm=na.rm)
		
		rxy.z <- (rxy.zc - rxz0.zc*ryz0.zc)/( sqrt(1-rxz0.zc^2)*sqrt(1-ryz0.zc^2) )
		return(rxy.z)
	}			
}	



```

\newpage

## Load data

```{r}
dat<-read.csv2("dat.no.miss.csv",stringsAsFactors = F)
```

### Variable transformations

#### Country

```{r}
table(dat$cntry)
```

#### Voting group

```{r}
#make voting group variable names unique to each country
dat$voting.group<-paste0(dat$cntry,": ",dat$vote.group.combined)

```

#### Centering Attitudes towards the Environment

```{r}
#recode the first variable to represent this attitude
dat$environ<-dat$inctxff.R
describe(dat$environ,fast=T)

#grand mean center
dat$environ.gmc<-dat$environ-mean(dat$environ,na.rm=T)

#obtain dataframe with country means and add to data

environ.cntry<-dat %>%
  group_by(cntry) %>%
  summarize(environ.cntry=mean(environ.gmc,na.rm=T))

dat<-left_join(x=dat,
               y=environ.cntry,
               by=c("cntry"))

#center individuals around country means

dat$environ.cntrymc<-dat$environ.gmc-dat$environ.cntry

#obtain dataframe with voting group means and add to data

environ.voting.group<-dat %>%
  group_by(voting.group) %>%
  summarize(environ.voting.group=mean(environ.cntrymc,na.rm=T))

dat<-left_join(x=dat,
               y=environ.voting.group,
               by=c("voting.group"))

#center individuals around voting group means

dat$environ.vgmc<-dat$environ.cntrymc-dat$environ.voting.group

#describe the variable

describe(dat$environ.vgmc,fast=T)

#rename as lvl1, lvl2, and lvl3

dat$environ.lvl1<-dat$environ.vgmc
dat$environ.lvl2<-dat$environ.voting.group
dat$environ.lvl3<-dat$environ.cntry
```

\newpage

#### Centering Political Engagement

```{r}
#correlation between the variables

corr.test(dat$nwspol.4,dat$polintr.R,adjust="none")

#rename the variable

dat$engagement<-dat$polint.agg

#descriptive statistics
psych::describe(dat$engagement,fast=T)

#grand mean center
dat$engagement.gmc<-dat$engagement-mean(dat$engagement,na.rm=T)

#obtain dataframe with country means and add to data

engagement.cntry<-dat %>%
  group_by(cntry) %>%
  summarize(engagement.cntry=mean(engagement.gmc,na.rm=T))

dat<-left_join(x=dat,
               y=engagement.cntry,
               by=c("cntry"))

#center individuals around country means

dat$engagement.cntrymc<-dat$engagement.gmc-dat$engagement.cntry

#obtain dataframe with voting group means and add to data

engagement.voting.group<-dat %>%
  group_by(voting.group) %>%
  summarize(engagement.voting.group=mean(engagement.cntrymc,na.rm=T))

dat<-left_join(x=dat,
               y=engagement.voting.group,
               by=c("voting.group"))

#center individuals around voting group means

dat$engagement.vgmc<-dat$engagement.cntrymc-dat$engagement.voting.group

#describe the centered variable

describe(dat$engagement.vgmc,fast=T)

#rename as lvl1, lvl2, and lvl3

dat$engagement.lvl1<-dat$engagement.vgmc
dat$engagement.lvl2<-dat$engagement.voting.group
dat$engagement.lvl3<-dat$engagement.cntry
```

\newpage

#### Rename and grand mean center the Attitudes towards refugees (pro-refugee attitudes indicate high scores)

```{r}
#grand mean center
dat$refugees<-dat$gvrfgap.R-mean(dat$gvrfgap.R,na.rm=T)
describe(dat$refugees,fast=T)

#rename
dat$refugees.gmc<-dat$refugees

#obtain dataframe with country means and add to data

refugees.cntry<-dat %>%
  group_by(cntry) %>%
  summarize(refugees.cntry=mean(refugees.gmc,na.rm=T))

dat<-left_join(x=dat,
               y=refugees.cntry,
               by=c("cntry"))

#center individuals around country means

dat$refugees.cntrymc<-dat$refugees.gmc-dat$refugees.cntry

#obtain dataframe with voting group means and add to data

refugees.voting.group<-dat %>%
  group_by(voting.group) %>%
  summarize(refugees.voting.group=mean(refugees.cntrymc,na.rm=T))

dat<-left_join(x=dat,
               y=refugees.voting.group,
               by=c("voting.group"))

#center individuals around voting group means

dat$refugees.vgmc<-dat$refugees.cntrymc-dat$refugees.voting.group

#describe the variable

describe(dat$refugees.vgmc,fast=T)

#rename as lvl1, lvl2, and lvl3

dat$refugees.lvl1<-dat$refugees.vgmc
dat$refugees.lvl2<-dat$refugees.voting.group
dat$refugees.lvl3<-dat$refugees.cntry
```

\newpage

#### Rename and Center the covariates aroung grand mean or logical middle points if applicable

```{r}
#grand-mean center age
dat$age<-dat$agea-mean(dat$agea,na.rm=T)
#sex around zero
dat$gender<-dat$gndr-1.5 #-0.5 males, 0.5 females
#rename occupation variable
dat$occup<-dat$isco.13
#grand-mean center education years
dat$educ<-dat$eduyrs-mean(dat$eduyrs,na.rm=T)
#residence around zero
dat$resid<-dat$rural-0.5 #-0.5 urban, 0.5 rural


```

\newpage

#### Voting group dummy-coded variables

```{r}

#recode if the party voted is =1, or not =0 anti-immigration
dat$anti.imm.party.dummy<-ifelse(is.na(dat$anti.imm.party.rule2),0,1)
#recode if the party voted is =1, or not =0 pro-environment
dat$pro.env.party.dummy<-ifelse(is.na(dat$pro.env.party.manual),0,1)

#dat$other.party.dummy<-ifelse(grepl("Other",dat$vote.group.combined),1,0)

#dummy-code not voting
dat$did.not.vote.dummy<-ifelse(grepl("did not vote",dat$vote.group.combined),1,0)
table(dat$did.not.vote.dummy)

#dummy-code "don't know"
dat$dont.know.dummy<-ifelse(grepl("Don't know",dat$vote.group.combined),1,0)
table(dat$dont.know.dummy)

#dummy-code invalid vote
dat$invalid.vote.dummy<-ifelse(grepl("Invalid vote",dat$vote.group.combined),1,0)
table(dat$invalid.vote.dummy)

#dummy-code "no answer"
dat$no.answer.dummy<-ifelse(grepl("No answer",dat$vote.group.combined),1,0)
table(dat$no.answer.dummy)

#dummy-code not-eligible: age
dat$not.eligible.age.dummy<-ifelse(grepl("not eligible: age",dat$vote.group.combined),1,0)
table(dat$not.eligible.age.dummy)

#dummy code not-eligible: citizenship
dat$not.eligible.citizenship.dummy<-ifelse(grepl("not eligible: citizenship",dat$vote.group.combined),1,0)
table(dat$not.eligible.citizenship.dummy)

#dummy-code not-eligible: other reasons
dat$not.eligible.other.dummy<-ifelse(grepl("not eligible: other",dat$vote.group.combined),1,0)
table(dat$not.eligible.other.dummy)

#add dummy-variable for other_party voting

dat<- dat %>%
  mutate(other.party.dummy:=case_when(
    anti.imm.party.dummy==1 |
      pro.env.party.dummy==1 |
      did.not.vote.dummy==1 |
      dont.know.dummy==1 |
      invalid.vote.dummy==1 |
      no.answer.dummy==1 |
      not.eligible.age.dummy==1 |
      not.eligible.citizenship.dummy==1 |
      not.eligible.other.dummy==1 ~0,
    TRUE~1
  ))

table(dat$other.party.dummy)

#recode the names for a new multi-category variable: all.parties.lvl2

dat<-dat %>%
  mutate(all.parties.lvl2:=case_when(
    did.not.vote.dummy==1~"Did not vote",
    dont.know.dummy==1~"Don't know",
    no.answer.dummy==1~"No answer",
    invalid.vote.dummy==1~"Invalid vote",
    not.eligible.age.dummy==1~"NE age",
    not.eligible.citizenship.dummy==1~"NE citizen",
    not.eligible.other.dummy==1~"NE other",
    other.party.dummy==1~"Other party",
    anti.imm.party.dummy==1~"Anti-immigration party",
    pro.env.party.dummy==1~"Pro-environment party",
  ),
  party:=case_when(
    other.party.dummy==1~"Other party",
    anti.imm.party.dummy==1~"Anti-immigration party",
    pro.env.party.dummy==1~"Pro-environment party",
    TRUE~NA_character_
  ))


```


#### Omit missing variables

```{r}
#missing values per each row
dat$analysis.miss<-
  is.na(dat$cntry)+
  is.na(dat$voting.group)+
  is.na(dat$refugees)+
  is.na(dat$environ)+
  is.na(dat$vote.group.combined)+
  is.na(dat$age)+
  is.na(dat$gender)+
  is.na(dat$occup)+
  is.na(dat$educ)+
  is.na(dat$resid)+
  is.na(dat$engagement)
table(dat$analysis.miss)

#include only those without any missing values
dat<-dat %>%
  filter(analysis.miss ==0)
```
