---
title: 'Exploratory analysis: covariate moderation'
output: 
  html_document: 
    keep_md: yes
    toc: yes
    toc_depth: 5
---

# Preparations

```{r child = 'Preparations.Rmd'}
```


# Exploratory analyses for moderators

## Does the association vary by age?

### Center the age variable

```{r}

describe(dat$age)

#already grand-mean centered

#obtain dataframe with country means and add to data

age.cntry<-dat %>%
  group_by(cntry) %>%
  summarize(age.cntry=mean(age,na.rm=T))

dat<-left_join(x=dat,
               y=age.cntry,
               by=c("cntry"))

#center individuals around country means

dat$age.cntrymc<-dat$age-dat$age.cntry

#obtain dataframe with voting group means and add to data

age.voting.group<-dat %>%
  group_by(voting.group) %>%
  summarize(age.voting.group=mean(age.cntrymc,na.rm=T))

dat<-left_join(x=dat,
               y=age.voting.group,
               by=c("voting.group"))

#center individuals around voting group means

dat$age.vgmc<-dat$age.cntrymc-dat$age.voting.group

#describe the variable

describe(dat$age.vgmc)

#rename as lvl1, lvl2, and lvl3

dat$age.lvl1<-dat$age.vgmc
dat$age.lvl2<-dat$age.voting.group
dat$age.lvl3<-dat$age.cntry
```

\newpage

### Model 1 (Same as H1 selected model but with centered age)

* Divide age variable by 10 to give interpretation by a decade

```{r}

dat$age.lvl1.10<-dat$age.lvl1/10

EX3.mod1<-lmer(refugees~(environ.lvl1|voting.group)+
                (0+environ.lvl1|cntry)+
                gender+occup+educ+resid+
                 age.lvl1.10+
                environ.lvl1,data=dat,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))



isSingular(EX3.mod1)
(VC.EX3.mod1<-getVC(EX3.mod1))
getFE(EX3.mod1)
```


\newpage

### Model 2 (interaction between age and environmental attitudes)

```{r}
EX3.mod2<-lmer(refugees~(environ.lvl1|voting.group)+
                (0+environ.lvl1|cntry)+
                gender+occup+educ+resid+
                 age.lvl1.10+
                environ.lvl1+
                 age.lvl1.10:environ.lvl1,
                 data=dat,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))


anova(EX3.mod1,EX3.mod2)
isSingular(EX3.mod2)
(VC.EX3.mod2<-getVC(EX3.mod2))
getFE(EX3.mod2)


```


\newpage

#### Marginal effects for ages at -1SD and +1SD

```{r message=FALSE}

EX3.mod2.trends<-
  emtrends(EX3.mod2,specs = c("age.lvl1.10"),var=c("environ.lvl1"),
           at=list(age.lvl1.10=c(
             
             mean(dat$age.lvl1.10)-sd(dat$age.lvl1.10),
             mean(dat$age.lvl1.10),
             mean(dat$age.lvl1.10)+sd(dat$age.lvl1.10)
             )))

(EX3.mod2.trends.tab<-data.frame(EX3.mod2.trends))


EX3.mod2.trends.tab$p<-
  2*(1-pnorm(abs(EX3.mod2.trends.tab$environ.lvl1.trend/
                   EX3.mod2.trends.tab$SE)))
EX3.mod2.trends.tab$adj.p<-
  p.adjust(EX3.mod2.trends.tab$p,method="holm")

EX3.mod2.trends.tab<-
  cbind(group=round(EX3.mod2.trends.tab[,1],2),
      round(EX3.mod2.trends.tab[,c(2,3)],2),
      round(EX3.mod2.trends.tab[,c(7,8)],4),
      round(EX3.mod2.trends.tab[,c(5,6)],2))
EX3.mod2.trends.tab

pairs(EX3.mod2.trends,adjust="none")

```

\newpage

## Does the association vary by sex?

### Center the sex variable

```{r}

describe(dat$gender)

#grand mean center
dat$gender.gmc<-dat$gender-mean(dat$gender,na.rm=T)

#obtain dataframe with country means and add to data

gender.cntry<-dat %>%
  group_by(cntry) %>%
  summarize(gender.cntry=mean(gender.gmc,na.rm=T))

dat<-left_join(x=dat,
               y=gender.cntry,
               by=c("cntry"))

#center individuals around country means

dat$gender.cntrymc<-dat$gender.gmc-dat$gender.cntry

#obtain dataframe with voting group means and add to data

gender.voting.group<-dat %>%
  group_by(voting.group) %>%
  summarize(gender.voting.group=mean(gender.cntrymc,na.rm=T))

dat<-left_join(x=dat,
               y=gender.voting.group,
               by=c("voting.group"))

#center individuals around voting group means

dat$gender.vgmc<-dat$gender.cntrymc-dat$gender.voting.group

#describe the variable

describe(dat$gender.vgmc)

#rename as lvl1, lvl2, and lvl3

dat$gender.lvl1<-dat$gender.vgmc
dat$gender.lvl2<-dat$gender.voting.group
dat$gender.lvl3<-dat$gender.cntry
```

\newpage

### Model 1 (Same as H1 selected model but with centered sex)

```{r}

EX1.mod1<-lmer(refugees~(environ.lvl1|voting.group)+
                (0+environ.lvl1|cntry)+
                age+occup+educ+resid+
                 gender.lvl1+
                environ.lvl1,data=dat,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))



isSingular(EX1.mod1)
getVC(EX1.mod1)
getFE(EX1.mod1)

```

\newpage

### Model 2 (interaction between sex and environmental attitudes)

```{r}
EX1.mod2<-lmer(refugees~(environ.lvl1|voting.group)+
                (0+environ.lvl1|cntry)+
                age+occup+educ+resid+
                 gender.lvl1+
                environ.lvl1+
                 gender.lvl1:environ.lvl1,
                 data=dat,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))


anova(EX1.mod1,EX1.mod2)
isSingular(EX1.mod2)
getVC(EX1.mod2)
getFE(EX1.mod2)
```

\newpage


## Does the association vary by education (years)?

### Center the education variable

```{r}

describe(dat$educ)

#already grand-mean centered

#obtain dataframe with country means and add to data

educ.cntry<-dat %>%
  group_by(cntry) %>%
  summarize(educ.cntry=mean(educ,na.rm=T))

dat<-left_join(x=dat,
               y=educ.cntry,
               by=c("cntry"))

#center individuals around country means

dat$educ.cntrymc<-dat$educ-dat$educ.cntry

#obtain dataframe with voting group means and add to data

educ.voting.group<-dat %>%
  group_by(voting.group) %>%
  summarize(educ.voting.group=mean(educ.cntrymc,na.rm=T))

dat<-left_join(x=dat,
               y=educ.voting.group,
               by=c("voting.group"))

#center individuals around voting group means

dat$educ.vgmc<-dat$educ.cntrymc-dat$educ.voting.group

#describe the variable

describe(dat$educ.vgmc)

#rename as lvl1, lvl2, and lvl3

dat$educ.lvl1<-dat$educ.vgmc
dat$educ.lvl2<-dat$educ.voting.group
dat$educ.lvl3<-dat$educ.cntry
```

\newpage

### Model 1 (Same as H1 selected model but with centered education)

```{r}

EX4.mod1<-lmer(refugees~(environ.lvl1|voting.group)+
                (0+environ.lvl1|cntry)+
                gender+occup+age+resid+
                 educ.lvl1+
                environ.lvl1,data=dat,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))



isSingular(EX4.mod1)
(VC.EX4.mod1<-getVC(EX4.mod1))
getFE(EX4.mod1)
```

\newpage

### Model 2 (interaction between education and environment attitudes)

```{r}

EX4.mod2<-lmer(refugees~(environ.lvl1|voting.group)+
                (0+environ.lvl1|cntry)+
                gender+occup+age+resid+
                 educ.lvl1+
                environ.lvl1+
                 environ.lvl1:educ.lvl1,data=dat,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))

anova(EX4.mod1,EX4.mod2)
isSingular(EX4.mod2)
(VC.EX4.mod2<-getVC(EX4.mod2))
getFE(EX4.mod2)


```


\newpage

#### Marginal effects for different levels of education

```{r message=FALSE}
EX4.mod2.trends<-
  emtrends(EX4.mod2,specs = c("educ.lvl1"),var=c("environ.lvl1"),
           at=
             list(educ.lvl1=
                    c(mean(dat$educ.lvl1)-sd(dat$educ.lvl1),
                      mean(dat$educ.lvl1),                              mean(dat$educ.lvl1)+sd(dat$educ.lvl1))))
(EX4.mod2.trends.tab<-data.frame(EX4.mod2.trends))


EX4.mod2.trends.tab$p<-
  2*(1-pnorm(abs(EX4.mod2.trends.tab$environ.lvl1.trend/
                   EX4.mod2.trends.tab$SE)))
EX4.mod2.trends.tab$adj.p<-
  p.adjust(EX4.mod2.trends.tab$p,method="holm")

EX4.mod2.trends.tab<-
  cbind(group=round(EX4.mod2.trends.tab[,1],2),
        round(EX4.mod2.trends.tab[,c(2,3)],2),
        round(EX4.mod2.trends.tab[,c(7,8)],4),
        round(EX4.mod2.trends.tab[,c(5,6)],2))
EX4.mod2.trends.tab


```

\newpage

## Does the association vary by place of residence (urban/rural)?

### Center the residence variable

```{r}

describe(dat$resid)
dat$resid.gmc<-dat$resid-mean(dat$resid,na.rm=T)


#obtain dataframe with country means and add to data

resid.cntry<-dat %>%
  group_by(cntry) %>%
  summarize(resid.cntry=mean(resid.gmc,na.rm=T))

dat<-left_join(x=dat,
               y=resid.cntry,
               by=c("cntry"))

#center individuals around country means

dat$resid.cntrymc<-dat$resid.gmc-dat$resid.cntry

#obtain dataframe with voting group means and add to data

resid.voting.group<-dat %>%
  group_by(voting.group) %>%
  summarize(resid.voting.group=mean(resid.cntrymc,na.rm=T))

dat<-left_join(x=dat,
               y=resid.voting.group,
               by=c("voting.group"))

#center individuals around voting group means

dat$resid.vgmc<-dat$resid.cntrymc-dat$resid.voting.group

#describe the variable

describe(dat$resid.vgmc)

#rename as lvl1, lvl2, and lvl3

dat$resid.lvl1<-dat$resid.vgmc
dat$resid.lvl2<-dat$resid.voting.group
dat$resid.lvl3<-dat$resid.cntry
```

\newpage

### Model 1 (Same as H1 selected model but with centered residence)

```{r}

EX5.mod1<-lmer(refugees~(environ.lvl1|voting.group)+
                (0+environ.lvl1|cntry)+
                gender+occup+age+educ+
                 resid.lvl1+
                environ.lvl1,data=dat,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))



isSingular(EX5.mod1)
(VC.EX5.mod1<-getVC(EX5.mod1))
getFE(EX5.mod1)
```

\newpage

### Model 2 (interaction between residence and environment attitudes)

```{r}

EX5.mod2<-lmer(refugees~(environ.lvl1|voting.group)+
                (0+environ.lvl1|cntry)+
                gender+occup+age+educ+
                 resid.lvl1+
                environ.lvl1+
                 environ.lvl1:resid.lvl1,data=dat,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))

anova(EX5.mod1,EX5.mod2)
isSingular(EX5.mod2)
(VC.EX5.mod2<-getVC(EX5.mod2))
getFE(EX5.mod2)




```


\newpage

## Does the association vary by occupational groups

### Model 1 (Same as H1 selected model)

```{r}

EX2.mod1<-lmer(refugees~(environ.lvl1|voting.group)+
                (0+environ.lvl1|cntry)+
                age+gender+educ+resid+occup+
                environ.lvl1,data=dat,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))



isSingular(EX2.mod1)
getVC(EX2.mod1)

```

\newpage


### Model 2 (Interaction between environment and occupational groups)

```{r}

EX2.mod2<-lmer(refugees~(environ.lvl1|voting.group)+
                (0+environ.lvl1|cntry)+
                age+gender+educ+resid+occup+
                environ.lvl1+
                 occup:environ.lvl1,data=dat,REML=F,
                 control=lmerControl(optimizer="bobyqa",
                                     optCtrl=list(maxfun=2e8)))


anova(EX2.mod1,EX2.mod2)
isSingular(EX2.mod2)
getVC(EX2.mod2)

```

\newpage

#### Marginal effects for each occupation group

```{r message=FALSE}
EX2.mod2.trends<-emtrends(EX2.mod2,specs = c("occup"),var=c("environ.lvl1"))
(EX2.mod2.trends.tab<-data.frame(EX2.mod2.trends))


EX2.mod2.trends.tab$p<-
  2*(1-pnorm(abs(EX2.mod2.trends.tab$environ.lvl1.trend/
                   EX2.mod2.trends.tab$SE)))
EX2.mod2.trends.tab$adj.p<-
  p.adjust(EX2.mod2.trends.tab$p,method="holm")

EX2.mod2.trends.tab<-
  cbind(group=EX2.mod2.trends.tab[,1],
        round(EX2.mod2.trends.tab[,c(2,3)],2),
        round(EX2.mod2.trends.tab[,c(7,8)],4),
        round(EX2.mod2.trends.tab[,c(5,6)],2))
EX2.mod2.trends.tab

#contrast for all groups against mean of other groups
contrast(EX2.mod2.trends, "del.eff", by = NULL,adjust=c("holm"))

```




